Quantitative Benchmarking for Automatically Generated Conjectures

### 2017-09-07 TEBenchmarkPaper/article.tex
More fixing up

\maketitle
IGNORE ISABELLE THEORIES FOR NOW!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
IGNORE CLUSTERING FOR NOW!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

intro (motivation, high-level background, eval is a problem),
background,
our proposal,
application (evaluation),
discussion (e.g. philosophical bits),
future work, conclusions, etc.

COMPARE COLTON ET AL.

Evaluation of conjectures, sets of conjectures, systems
what constitutes a good/desired output (in each author's terms)? This combines
their evaluation with their (system-specific) assumptions (e.g. IsaCoSy doesn't
generate redundant terms: non-redundancy is clearly a goal, but it's not measured
in eval since it's assumed based on system's construction)

Look carefully at each system's (paper's) eval section
Make a table like tabl 1 of Colton et al

Make it clear that methodology is the point (it's not just a detail, like in a
chemistry paper for example)

@@

% TODO: Preaching to the choir;
% Jumping around a bit, rephrase
% Look at phd symposium 2017, moa's conjecture synthesis paper, etc.
% We can just jump straight in to automated conjecture generation

% Maybe lay out argument in bullet points; flesh out sentences later
Automated \emph{theory exploration} (ATE), also known as
\emph{conjecture generation/formation}, is the open-ended problem of producing
conjectures about a given logical theory which are somehow ``interesting''.

@@

% Lay out the field of TE: motivation, systems, evaluation, problems...
% Compare to

The field of mathematics has been transformed by the introduction of computers:
calculations are now routinely automated, allowing humans to focus their efforts
on higher-level, more creative activities like proving theorems; \emph{automated
theorem proving} (ATP) seeks to automate this process too, and was one of the
first tasks studied in the field of artificial intelligence
\cite{newell1956logic,sutcliffe2001evaluating}. Despite this effort, ATP
methods are not widely used by traditional mathematicians, although they have
seen use in applied areas such as hardware and software verification
\cite{Moore:2003}.

% FIXME: Be more specific about our contributions

Even if ATP became widespread, the automation of theorem proving would still
leave an important aspect of mathematics inaccessible to computers: conjecturing
those theorems (and the definitions they involve) in the first place. Automated
methods for posing good conjectures have obvious applications for those who must
invent a constant stream of novel questions, e.g. for examinations or research.
A larger impact may be found in areas such as computer programming, security and
verification, where the benefits of formal methods are hindered by the
difficulty of their application; tools to aid this process, e.g. by proposing
test cases, can reduce this cost.

Many automated systems and algorithms have been developed to generate
conjectures, but it is difficult to judge or compare their performance in a
quantitative way. There are many aspects to consider when judging the quality of
a conjecture: how likely it is to be true, how difficult it is to prove, its
implications, and so on.

@@

One problem with this approach is the small size of these libraries. For
example, a benchmark based on Isabelle/HOL's theory of natural numbers contains
only 4 definitions and 12 theorems. Whilst such benchmarks allow
comparison between different approaches, their narrow scope doesn't provide much
indication of performance in different, especially \emph{novel}, domains.

@@

\section{Existing Evaluations}
\label{sec:previous}

Theory exploration tools such as
IsaCoSy~\cite{Johansson.Dixon.Bundy:conjecture-generation}, QuickSpec and
IsaScheme have been evaluated using the standard library of the Isabelle theorem
prover as their ground truth.

ALGORITHM:

Concatenate all definitions from all files together.
For each definition, prefix its name with the filename it came from and update
all references.
Normalise all local variable names.
Remove any identical definitions, leaving the first one in place and updating
any references to the removed definitions to use this one.
Repeat, until no further removals are possible.

\section{Evaluation}
\label{sec:evaluation}

@@

\section{Conclusion}
\label{sec:conclusion}

>>>>

\maketitle
%IGNORE ISABELLE THEORIES FOR NOW!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
%IGNORE CLUSTERING FOR NOW!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

%COMPARE COLTON ET AL.

% Evaluation of conjectures, sets of conjectures, systems
% what constitutes a good/desired output (in each author's terms)? This combines
% their evaluation with their (system-specific) assumptions (e.g. IsaCoSy doesn't
% generate redundant terms: non-redundancy is clearly a goal, but it's not measured
% in eval since it's assumed based on system's construction)

% Look carefully at each system's (paper's) eval section
% Make a table like tabl 1 of Colton et al

% Make it clear that methodology is the point (it's not just a detail, like in a
% chemistry paper for example)

@@

% Lay out the field of TE: motivation, systems, evaluation, problems...

Automated \emph{theory exploration} (ATE), also known as
\emph{conjecture generation/formation}, is the open-ended problem of producing
conjectures about a given logical theory which are somehow ``interesting''.

@@

%- Existing approaches are difficult to compare
% - They want different things
% - They're evaluated differently
% - They're evaluated ``locally'' (as white boxen)

% Look at phd symposium 2017, moa's conjecture synthesis paper, etc.
% We can just jump straight in to automated conjecture generation

\section{Background}
\label{section:background}

\begin{figure}
  \label{figure:list_theory}
  \begin{equation*}
    Nil : \forall (a : Type). List a
    Cons : \forall (a : Type). \forall (x : a). \forall (xs : List a). List a

    Z : Nat
    S : \forall (p : Nat). Nat

    head(Cons(x, xs)) = x
    tail(Cons(x, xs)) = xs
    append(xs, Nil) = xs
    append(Cons(x, xs), ys) = Cons x (append xs ys)
    reverse(Nil) = Nil
    reverse(Cons(x, xs)) = append(reverse(xs), Cons(x, Nil))
    length(Nil) = Z
    length(Cons(x, xs)) = S (length(xs))
    map(f, Nil) = Nil
    map(f, Cons(x, xs)) = Cons(f(x), map(f, xs))
    foldl(f, x, Nil) = x
    foldl(f, x, Cons(y, ys)) = foldl(f, f(x, y), xs)
    foldr(f, Nil, y) = y
    foldr(f, Cons(x, xs), y) = f(x, foldr(f, xs, y))
  \end{equation*}
  \caption{A simple theory defining $Nat$ and $List$ types and some associated
    list operations, taken from \cite{Johansson.Dixon.Bundy:conjecture-generation}}
\end{figure}

If we have a logical theory, for example the theory of lists shown in figure
\ref{figure:list_theory} Many automated systems and algorithms have been developed to generate
conjectures, but it is difficult to judge or compare their performance in a
quantitative way. We consider three distinct aspects to evaluate:

\begin{itemize}
\item Each conjecture, to judge how ``interesting'' it is.
\item The \emph{set} of conjectures produced, to assess things like consistency
  and redundancy.
\item Performance of the system, to find the balance struck between output
  quality and time taken.
\end{itemize}

There are many aspects to consider when judging the quality of
a conjecture: how likely it is to be true, how difficult it is to prove, its
implications, and so on.

@@

One problem with this approach is the small size of these libraries. For
example, the benchmark based on Isabelle/HOL's theory of natural numbers given
in~\cite{Johansson.Dixon.Bundy:conjecture-generation} contains
only 4 definitions and 12 theorems. Whilst such benchmarks allow
comparison between different approaches, their narrow scope doesn't provide much
indication of performance in different, especially \emph{novel}, domains.

\section{Our Proposal}
\label{section:proposal}

@@

%ALGORITHM:

%Concatenate all definitions from all files together.
%For each definition, prefix its name with the filename it came from and update
%all references.
%Normalise all local variable names.
%Remove any identical definitions, leaving the first one in place and updating
%any references to the removed definitions to use this one.
%Repeat, until no further removals are possible.

\section{Application}
\label{section:application}

% evaluation

@@

\section{Discussion}
\label{section:application}

% (e.g. philosophical bits),

\section{Conclusion}
\label{section:conclusion}

% future work, etc.
